---
title: "Finacial risk volatility modeling (ARCH, GARCH, & EGRCH) with R"
author: "Vijaya Gagrani"
date: "May 2015"
output:
  slidy_presentation: default
  ioslides_presentation:
  keep_md: yes
  transition: faster
  widescreen: yes
  beamer_presentation: default
---
   
```{r Set_ChunkOptions, echo=FALSE}
#Select Chunk and package options @ http://yihui.name/knitr/options
knitr::opts_chunk$set(comment = NA, echo= FALSE, message = FALSE, fig.align='center', warning = FALSE,cache=FALSE)
```
## Background and Objective

Finacial risk univariate volatility models included in this analyis are ARCH, GARCH, and EGRCH. These models are based on a least squares models assumptions that the expected value of all squared error terms or variance is same at any given point (homoskendasticity). The opposite concept is heterskendasticity, where variance of the error term is not constant for all points, and thats what the center of the Autoregressive Conditional Heteroskedasticity models (ARCH).

Volatility is important in assessing financial risk, portfolio selection, and derivative pricing. However, in financial time series, volatility occurs in clusters like periods of high volatility follows with high volatility and vice versa.  In addtion, for financial time series more recent events would be more relevant then the distant past events.  The ARCH models utilize the heteroskedasticity in time series data to measure volatility (standard deviation). These models are applied to forecast mean and variance of return based on the past information. In addtion, these models allow to use the best weights in forecasting variance instead of using equal weights for all past events.

An ARCH model is simila to the AR (1) model on squared residuals. In an ARCH (1) model, next period’s variance only depends on last period’s squared residual. The ARCH model is extended as the Generalized ARCH or GARCH model by making it conditional to both squared residuals and squared observations in past. The typical GARCH model is described as GARCH(1,1) model. The first subscript in GARCH reefers to autoregressive lags or ARCH and second subscript refers to the moving average lags, respectively. These models are basically setup for one period forecast, however a long-horizon forecasting can be constructed. The ARCH/GARCH approaches are widely used in situations where volatility of return is a central issue.  


Financial institutions measure the risk using a Value-at-Risk (VaR), usually calculated at the 99% confidence level. Generally, for a stock prices the low volatility periods follows by the periods of high volatility and vice versa (cluster volatility). This phenomenon is visually detectable in simple line plots. In addtion, we can test it by applying the two commonly used tests(The Ljung-Box test for autocorrelation in squared returns and the Lagrange Multiplier (LM test)). In rugarch library there are diffrent functions for model specification, parameter estimation, backtesting, and forecasting. 

```{r}
Box.test(coredata(intc^2), type = "Ljung-Box", lag = 12)
#Reject the null hypothesis of no autocorrelations in the squared returns at the 1% significance level
#LM test
install.packages("FinTS")
library("FinTS")
ArchTest(coredata(intc))

library("rugarch")
#specify the model
intc_garch11_spec <- ugarchspec(variance.model = list(+   garchOrder = c(1, 1)),+  mean.model = list(armaOrder = c(0, 0)))
#model estimation
intc_garch11_fit <- ugarchfit(spec = intc_garch11_spec,+  data = intc)
#backtesting: to check the model perfromance 
intc_garch11_roll <- ugarchroll(intc_garch11_spec, intc,+   n.start = 120, refit.every = 1, refit.window = "moving",+   solver = "hybrid", calculate.VaR = TRUE, VaR.alpha = 0.01,+   keep.coef = TRUE)
#examining backtest report
report(intc_garch11_roll, type = "VaR", VaR.alpha = 0.01,+   conf.level = 0.99)
#Kupiec's unconditional coverage compares the number of expected versus actual exceedances given the tail probability of VaR, while the Christoffersen test is a joint test of the unconditional coverage and the independence of the exceedances. In our case, despite the actual five exceedances versus an expectation of three, we can't reject the null hypothesis that the exceedances are correct and independent.
intc_VaR <- zoo(intc_garch11_roll@forecast$VaR[, 1])
index(intc_VaR) <- as.yearmon(rownames(intc_garch11_roll@forecast$VaR))
intc_actual <- zoo(intc_garch11_roll@forecast$VaR[, 2])
index(intc_actual) <- as.yearmon(rownames(intc_garch11_roll@forecast$VaR))
plot(intc_actual, type = "b", main = "99% 1 Month VaR Backtesting",+   xlab = "Date", ylab = "Return/VaR in percent")
lines(intc_VaR, col = "red")
legend("topright", inset=.05, c("Intel return","VaR"), col = c("black","red"), lty = c(1,1))
#forecasting
intc_garch11_fcst <- ugarchforecast(intc_garch11_fit, n.ahead = 12)

```
## Vector Autoregressive Models (VAR) with R

The Vector autoregressive models (VAR) are the multivariate extensions of the univariate autoregressive (AR) models, captures the linear dependencies among multiple time series. As per the VAR model each variable is evolved through a linear function of all other lagged variables. All variables depends on the past values of its own as well as the other variables in consideration. For a VAR model, all variables have to be stationary or show the same order of integration. In case if the variables are nonstationary, or cointegrated, the error correction term is included in the VAR and the model is known as Vector Error Correction Model (VECM). 
A nonstationary time series have means, variances, and covariances that change over time. It occurs as trends, cycles, random walks, or combinations of the three. Regression models without accounting the nonstationarity results into spurious regressions. The cointegration technique is useful to investigate cointegrated or nonstationary variables. The Engle-Granger tests and the Johansen-Juselius tests allow testing the cointegration.
Augmented Dickey–Fuller (ADF) t-statistic test: small p-values suggest (library(urca))
A VAR model is primarily used to assess the Impulse-response from a stress. Essentially, an impulse-response function shows how a variable reacts (response) to a shock (impulse) influencing any other variable in the system. For example the impulse response of 1SD shock in S&P prices for a univariate analysis. For multivariate analysis, covariance matrix is used.
In this analysis following three datasets are used:
1. Equity Return (wfc)
2. Stock Index (S&P500)
3. US Treasury bond interest rates
The primary purpose is to make a forecast for the S&P index by using the additional variables and to identify impulse responses. The above three variable holds a long term relationship.


```{}
# data download, Interest rates are downloaded from the FRED (Federal Reserve Economic Data) data source. rest two are from yahoo finance.
#The R package, "vars" provide an excellent framework for this analysis.
# Important R packages for this analysis are vars and quantmod.

getSymbols('MSFT', from='2004-01-02', to='2014-03-31')
getSymbols('SNP', from='2004-01-02', to='2014-03-31')
getSymbols('DTB3', src='FRED')
plot...
#Reduced VAR model
var1 <- VAR(dataDaily, lag.max=4, ic="AIC")
VARselect(dataDaily,lag.max=4)
summary(var1)

var.pred <- predict(var1, n.ahead=10, ci=0.95)
var.irf <- irf(var1)
plot(var.irf)
#structural VAR

#SVAR

```

Berlinger, E., Illés, F., Badics, M., Banai, Á., Daróczi, G., Dömötör, B., ... & Vidovics-Dancs, Á. (2015). Mastering R for Quantitative Finance. Packt Publishing Ltd.

Fabozzi, F. J., Focardi, S. M., Rachev, S. T., & Arshanapalli, B. G. (2014). The Basics of Financial Econometrics: Tools, Concepts, and Asset Management Applications. John Wiley & Sons.

http://www.investopedia.com

http://www.cmat.edu.uy/~mordecki/hk/engle.pdf
http://faculty.washington.edu/ezivot/econ589/econ589univariateGarch.r
http://faculty.washington.edu/ezivot/econ589/univariateGarch2012powerpoint.pdf
http://www.r-bloggers.com/a-practical-introduction-to-garch-modeling/
https://onlinecourses.science.psu.edu/stat510/node/61
