---
title: "Vector or multivariate time series analysis with R"
author: "Vijaya Gagrani"
date: "May 2015"
output:
  slidy_presentation: default
  ioslides_presentation:
  keep_md: yes
  transition: faster
  widescreen: yes
  beamer_presentation: default
---
   
```{r Set_ChunkOptions, echo=FALSE}
#Select Chunk and package options @ http://yihui.name/knitr/options
knitr::opts_chunk$set(comment = NA, echo= FALSE, message = FALSE, fig.align='center', warning = FALSE,cache=FALSE)
```
## Background and Objective

```{r Initial_Settings_n_Basic_Stat}
#set working directory
#setwd("C:/Users/gagranis/Documents/Finance/Coursera/qfinance")
options(digits=4, width=70)
library("PerformanceAnalytics")
library("xts")
library("zoo")
library("tseries")
library("plyr")
library("mvtnorm")
library("splusTimeSeries")
library("boot")
library("FinTS") #  for LM test
library("rugarch")
```

## Vector Autoregressive Models (VAR) with R

This repository focuses to explore the econometric models and methods to analyze multiple finacial time series of returns, known as the vector or multivariate time series analysis models.

The Vector autoregressive models (VAR) are the multivariate extensions of the univariate autoregressive (AR) models, captures the linear dependencies among multiple time series. As per the VAR model each variable is evolved through a linear function of all other lagged variables. All variables depends on the past values of it's own as well as the other variables in consideration. For a VAR model, all variables have to be stationary, else show the same order of integration. In case if the variables are nonstationary, or cointegrated, the error correction term is included in the VAR and then the model is considered as the Vector Error Correction Model (VECM). 

In addition to model the linear dependencies on multiple varibles, VAR models are used to model the covariance of multiple assessts using the multivariate ARCH/GARCH model often known as the VEC-GARCH model (Vector Error Correction Models), replacing the volatility parameter with the covariance matrix.

Statistical properties of the stationary time series do not changes for a shift in time. In a strict stationary time series all moments of all degrees stays same. However, for general purpose a weak defination is used, where only two moments (mean and variance) unchaged for a shift in time (or mean reversion), and autocorrelation/covariance only depends on the lag length.

The stylized facts are statistical propertis which stays consitent over a time period for a type of a dataset. For finacial time series, returns often exhibits fat tails and autocorrelation. 

In the cross correlation matices(CCM) of a week stationary vector time sereis, the diagonal elements contains the autocorrrelation function, the off diagonal elements measures the concurrent linear relationship.

The non-stationary time series wanders arbitrarily far from the mean. Therefore, these time series shows the stocastic trends, cycles, random walks, or combinations of all. In order to find out wheather a time series is non-stationary, unit root tests are carried out using the autoregressive models, one of such test is the Dickey-Fuller test. However, if residuals are autocorrelated, use the Augumented Dickey-Fuller tests (ADF). The ADF test includes addtional lags. The unit root test are used to detrend data to render staionarity using first diffrenced or regressed on determistic functions of time or using the cointegration techniques for the long-run relations.   

Regression models without accounting the non-stationarity results into spurious regressions (situation like when two economic variables seems related over time, but in fact they are not). Cointegration technique is useful to investigate the cointegrated or nonstationary variables. If all variables in an econometric model share a common stochastic trend, the cointegration analysis uncovers the long-term relationship and the short-term dynamics while controling the spurious regression problem. The following two tests: the Engle-Granger and the Johansen-Juselius are used to test the cointegration. 

(http://denizstij.blogspot.com/2013/11/cointegration-tests-adf-and-johansen.html
https://cran.r-project.org/web/packages/egcm/egcm.pdf)
Augmented Dickey–Fuller (ADF) t-statistic test: small p-values suggest (library(urca)).

A VAR model is primarily used to assess the Impulse-response from a stress. Essentially, an impulse-response function shows how a variable reacts (response) to a shock (impulse) influencing any other variable in the system. For example the impulse response of 1SD shock in S&P prices for an univariate analysis. 

In this analysis following three datasets are used:

1. Equity Return: (wfc) : Source: CRSP, "Center for Reserch in Security Prices"
2. Real labor income growth : log change in income from labor minus the log change in core inflation :FRED
3.  The term premium: it is the diffrence between the yield on a 10 year constant maturity bond and the 3-month t-bill rate: FRED
4.  The relative t-bill rate: current yeld on a 1-month t bill minus the average yield over the past 12 months :Ken French's web site
5.  the divident yield: the diffrence in returns on wfc with and without dividens :CRSP
6.  inflation and gdp

The primary purpose is to make a forecast for the S&P index by using two additional variables and to identify impulse responses. The above three variable holds a long term relationship.

The main advantage with SVAR analysis is that the necessary restrictions on the estimated reduced VAR model, required for identification of the underlying structural model, can be provided by economic theory.These restrictions can be either contemporaneous or long-run in nature depending on whether the underlying disturbances are considered to be temporary or permanent in nature. Once the identification is achieved it is possible to recover the structural shocks. These shocks can then be used to generate impulse response and variance decomposition functions to assess the dynamic impacts of different economic variables. In addition these functions can be used to test whether such shocks affect the economic variables as economic theory would predict so providing a check on the theory.

Granger Causality is a standard method to know if a varaible is useful in predicting other varible. Presence of GC indicates that a VAR instead of a univariate model is needed. 

Impulse Response Function(IRFs): In univariate time series, ACF tells about the shocks decay. However, for the vector data, it can affect the other varibles in consideration. The popular way is to simulate the effect of a unit shock. 

Cointegration: In situation when time series under consideration for VAR shows a temporary deviation from there long run relationship or trend, but ultimately mean reverting to this trend, the cointegration and error correction provide the tool to analyze temporary deviations from long run equlibria.

Error Correction Models(ECM): An ECM is a structure, which overns short deviations from the trend.

```{}
# data download, Interest rates are downloaded from the FRED (Federal Reserve Economic Data) data source. rest two are from yahoo finance.

#Board of Governors of the Federal Reserve System (US), 3-Month Treasury Bill: Secondary Market Rate [DTB3], retrieved from FRED, Federal Reserve Bank of St. Louis https://research.stlouisfed.org/fred2/series/DTB3/, November 7, 2015.


getSymbols("VBLTX", src = "yahoo", from ="1994-03-01", to = "2015-10-31")
getSymbols("wfc", src = "yahoo", from ="1994-03-01", to = "2015-10-31")
getSymbols("^GSPC",src = "yahoo", from ="1994-03-01", to = "2015-10-31")
getSymbols('DTB3', src='FRED')
DTB3<-DTB3["1994-03-01::2015-12-31"]

getSymbols('UNRATE',src='FRED')
chartSeries(UNRATE)
getSymbols("FEDFUNDS",src="FRED")
chartSeries(FEDFUNDS)
getSymbols("CPIAUCSL",src="FRED")
chartSeries(CPIAUCSL)
getSymbols("M1SL",src="FRED")
chartSeries(M1SL)
getSymbols("GDPDEF",src="FRED")
chartSeries(GDPDEF)
getSymbols("GDP",src="FRED")
chartSeries(GDP)
getSymbols("TB3MS",src="FRED")
chartSeries(TB3MS)
getSymbols("TWEXBMTH",src="FRED")
chartSeries(TWEXBMTH)

#Considerfocusingon3variables
ymat0<-merge(UNRATE,FEDFUNDS,CPIAUCSL)
par <- par("mar")
par(mar=c(1,1,1,1))
par(mfcol=c(3,1))
plot(ymat0[,1],main=dimnames(ymat0)[[2]][1])
plot(ymat0[,2],main=dimnames(ymat0)[[2]][2])
plot(ymat0[,3],main=dimnames(ymat0)[[2]][3])

# Extract window from 1960-2015

ymat0.0<-window(ymat0, start = as.Date("1960-01-01"),end = as.Date("2015-12-31"))

dim(ymat0.0)
head(ymat0.0)

acf(ymat0.0, lag.max=10)

#Vector Autoregressive (VAR) Model
ymat0.0.VAR.const<-VARselect(ymat0.0, lag.max=12, type="const")

ymat0.0.VAR.const$selection
ymat0.0.VAR.const.0<-VAR(ymat0.0,p=ymat0.0.VAR.const$selection[3],type="const")

summary(ymat0.0.VAR.const.0)


#Impulse Response Functions for a Fitted VAR(p) Model
plot(irf(ymat0.0.VAR.const.0, impulse="UNRATE"))

# Ordinary and Partial Autocorrelations of Differenced Series

ymat00.0<-na.omit(diff(ymat0.0))
par <- par("mar")
par(mar=c(1,1,1,1))
par(mfcol=c(3,1))
plot(ymat00.0[,1],main=dimnames(ymat00.0)[[2]][1])
plot(ymat00.0[,2],main=dimnames(ymat00.0)[[2]][2])
plot(ymat00.0[,3],main=dimnames(ymat00.0)[[2]][3])

acf(ymat00.0, lag.max=10)

#Vector Autoregressive (VAR) Model with Differenced Series
ymat00.0.VAR.const<-VARselect(ymat00.0, lag.max=12, type="const")
ymat00.0.VAR.const

ymat00.0.VAR.const.0<-VAR(ymat00.0,p=ymat00.0.VAR.const$selection[3],type="const")
options(show.signif.stars=FALSE)
summary(ymat00.0.VAR.const.0)

plot(irf(ymat00.0.VAR.const.0, impulse="UNRATE"))

```

Berlinger, E., Illés, F., Badics, M., Banai, Á., Daróczi, G., Dömötör, B., ... & Vidovics-Dancs, Á. (2015). Mastering R for Quantitative Finance. Packt Publishing Ltd.

Fabozzi, F. J., Focardi, S. M., Rachev, S. T., & Arshanapalli, B. G. (2014). The Basics of Financial Econometrics: Tools, Concepts, and Asset Management Applications. John Wiley & Sons.

http://www.investopedia.com

http://www.cmat.edu.uy/~mordecki/hk/engle.pdf
http://faculty.washington.edu/ezivot/econ589/univariateGarch2012powerpoint.pdf
http://www.r-bloggers.com/a-practical-introduction-to-garch-modeling/
https://onlinecourses.science.psu.edu/stat510/node/61
http://faculty.washington.edu/ezivot/econ584/notes/varModels.pdf
https://www.otexts.org/fpp/9/2
http://freakonometrics.hypotheses.org/12729 : unit root tests
http://www.ncbi.nlm.nih.gov/pmc/articles/PMC3223325/ : nih 
https://www.kevinsheppard.com/images/5/56/Chapter5.pdf :VAR pdf
https://www.youtube.com/watch?v=uBeM1FUk4Ps :  mit
http://dybfin.wustl.edu/teaching/finopt/pdfslides/eigen.pdf :egnevalue
https://cran.r-project.org/web/packages/vars/vignettes/vars.pdf :var example
http://ftp.uni-bayreuth.de/math/statlib/R/CRAN/doc/vignettes/vars/vars.pdf :var example2
https://onlinecourses.science.psu.edu/stat510/node/79 :Penn State

http://support.sas.com/documentation/cdl/en/etsug/63348/HTML/default/viewer.htm#etsug_varmax_sect048.htm :SAS example

http://ocw.mit.edu/courses/mathematics/18-s096-topics-in-mathematics-with-applications-in-finance-fall-2013/case-studies/MIT18_S096F13_CaseStudy5.pdf: macroeconomic VAR model
