
---
title: "Portfolio Theory"
author: "Vijaya Gagrani"
date: "May 2015"
output:
  slidy_presentation: default
  ioslides_presentation:
    keep_md: yes
    transition: faster
    widescreen: yes
  beamer_presentation: default
---

## Background and Objective

The expected returns(average) are the best guesses for the monthly returns, however returns are random varibles therefore realized returns may be diffrent from our expectations. Variance provides a measure towards uncertain monthly returns, hence considered as a risk measurment tool. All wealth is exhoused between asset A and B. Monthly return is a linear combination or weighted average of the two returns.http://faculty.washington.edu/ezivot/econ424/introductionPortfolioTheory.pdf.
http://faculty.washington.edu/ezivot/econ424/portfolioTheoryMatrix.pdf
http://faculty.washington.edu/ezivot/econ424/portfolio.r

```{r Set_ChunkOptions, echo=FALSE}
#Select Chunk and package options @ http://yihui.name/knitr/options
knitr::opts_chunk$set(comment = NA, echo= FALSE, message = FALSE, fig.align='center', warning = FALSE,cache=FALSE)

options(digits=4, width=70)
library("PerformanceAnalytics")
```

## 2 asset example

```{r}
mu_A = 0.175
sig_A = 0.258
sig2_A = sig_A^2
mu_B = 0.055
sig_B = 0.115
sig2_B = sig_B^2
rho_AB = -0.164
sig_AB = rho_AB*sig_A*sig_B
w0 = 1000
VaR_A = (mu_A + sig_A*qnorm(0.05))*w0
VaR_A
VaR_B = (mu_B + sig_B*qnorm(0.05))*w0
VaR_B
```

## example portfolios and VaR
```{r}
x_A = 0.5
x_B = 0.5
mu_p1 = x_A*mu_A + x_B*mu_B
sig2_p1 = x_A^2 * sig2_A + x_B^2 * sig2_B + 2*x_A*x_B*sig_AB
sig_p1 = sqrt(sig2_p1)
mu_p1
sig2_p1
sig_p1
VaR_p1 = (mu_p1 + sig_p1*qnorm(0.05))*w0
VaR_p1
# note: portfolio VaR is not a weighted average of individual asset VaR
x_A*VaR_A + x_B*VaR_B

```

### mean-variance efficient portfolios
Efficent portfolios have the highest expected return for a given level of risk.
Assumptions: Returns are covariance stationary and ergodic, jointly normally distrbuted means means, variances, and covariances of returns are constant over the investment horizon.For a given level of risk, combinations that are in the upper left conrner are the best (efficent portfolios) and those in the lower right corner(inefficent portfolios) are the worst portfolios. portfolio at the bottom of the parabola has the smallest variance among all fesiable protfolio and called global minimum variance portfolio. 


```{r}


x_A = seq(from=-0.4, to=1.4, by=0.1)
x_B = 1 - x_A
mu_p = x_A*mu_A + x_B*mu_B
sig2_p = x_A^2 * sig2_A + x_B^2 * sig2_B + 2*x_A*x_B*sig_AB
sig_p = sqrt(sig2_p)

# minimum variance portfolio
# global minmum protfolio: Lagrange multipliers

xA_min = (sig2_B - sig_AB)/(sig2_A + sig2_B - 2*sig_AB)
xB_min = 1 - xA_min
xA_min
xB_min
mu_p_min = xA_min*mu_A + xB_min*mu_B
sig2_p_min = xA_min^2 * sig2_A + xB_min^2 * sig2_B + 2*xA_min*xB_min*sig_AB
sig_p_min = sqrt(sig2_p_min)
mu_p_min
sig_p_min

# create portfolio plot
cex_val = 2
plot(sig_p, mu_p, type="b", pch=16, cex = cex_val,
     ylim=c(0, max(mu_p)), xlim=c(0, max(sig_p)),
     xlab=expression(sigma[p]), ylab=expression(mu[p]), cex_lab=cex_val,
     col=c(rep("red", 6), "blue", rep("green", 12)))
text(x=sig_A, y=mu_A, labels="Asset A", pos=4, cex = cex_val)
text(x=sig_B, y=mu_B, labels="Asset B", pos=4, cex = cex_val)
text(x=sig_p_min, y=mu_p_min, labels="Global min", pos=2, cex = cex_val)


# portfolio frontier with varying values of rho
# The shape of the investment possibilities set is senstive to the correlation between assessts.
rho_AB_vals = c(-0.9,-0.5, -0.25, 0, 0.25, 0.5, 0.9)
x_A = seq(from=-0.4, to=1.4, by=0.1)
x_B = 1 - x_A
mu_p = x_A*mu_A + x_B*mu_B
sig2_p = x_A^2 * sig2_A + x_B^2 * sig2_B + 2*x_A*x_B*rho_AB_vals[1]*sig_A*sig_B
sig_p = sqrt(sig2_p)

# create portfolio plot
cex_val = 1.5
plot(sig_p, mu_p, type="b", pch=16, cex = cex_val,
     ylim=c(0, max(mu_p)), xlim=c(0, max(sig_p)),
     xlab=expression(sigma[p]), ylab=expression(mu[p]),
     cex_lab = cex_val, col=1)
text(x=sig_A, y=mu_A, labels="Asset A", pos=4, cex = cex_val)
text(x=sig_B, y=mu_B, labels="Asset B", pos=4, cex = cex_val)

for (i in 2:length(rho_AB_vals)) {
  sig2_p = x_A^2 * sig2_A + x_B^2 * sig2_B + 2*x_A*x_B*rho_AB_vals[i]*sig_A*sig_B
  sig_p = sqrt(sig2_p)
  points(sig_p, mu_p, type="b", pch=16, col=i, cex=cex_val)
}
legend(x="bottomright", legend=paste("rho", as_character(rho_AB_vals), sep="="), 
       col=1:length(rho_AB_vals), lty=1, pch=16, cex = cex_val)

#
# show portfolio plot with rho = 1 and rho = -1
#
rho_AB = 1
sig_AB = rho_AB*sig_A*sig_B
mu_p_1 = x_A*mu_A + x_B*mu_B
sig2_p_1 = x_A^2 * sig2_A + x_B^2 * sig2_B + 2*x_A*x_B*sig_AB
sig_p_1 = sqrt(sig2_p_1)
rho_AB = -1
sig_AB = rho_AB*sig_A*sig_B
mu_p_m1 = x_A*mu_A + x_B*mu_B
sig2_p_m1 = x_A^2 * sig2_A + x_B^2 * sig2_B + 2*x_A*x_B*sig_AB
sig_p_m1 = sqrt(sig2_p_m1)

plot(sig_p_1, mu_p_1, type="b", pch=16, cex = cex_val,
     ylim=c(0, max(mu_p)), xlim=c(0, max(sig_p)),
     xlab=expression(sigma[p]), ylab=expression(mu[p]))
points(sig_p_m1, mu_p_m1, type="b", col="blue", pch=22, cex = cex_val)
text(x=sig_A, y=mu_A, labels="Asset A", pos=4, cex = cex_val)
text(x=sig_B, y=mu_B, labels="Asset B", pos=4, cex = cex_val)
legend(x="topleft", legend=c(expression(rho==1), expression(rho==-1)),
       col=c("black", "blue"), pch=c(16,22), cex = cex_val)

#
# show efficient portfolios
#
rho_AB = -0.164
sig_AB = rho_AB*sig_A*sig_B
x_A = seq(from=-0_4, to=1_4, by=0_1)
x_B = 1 - x_A
mu_p = x_A*mu_A + x_B*mu_B
sig2_p = x_A^2 * sig2_A + x_B^2 * sig2_B + 2*x_A*x_B*sig_AB
sig_p = sqrt(sig2_p)
plot(sig_p, mu_p, type="b", pch=16, ylim=c(0, max(mu_p)), cex = cex_val,
     xlim=c(0, max(sig_p)),
     xlab=expression(sigma[p]), ylab=expression(mu[p]),
     col=c(rep("red", 6), rep("green", 13)))
text(x=sig_A, y=mu_A, labels="Asset A", pos=4, cex = cex_val)
text(x=sig_B, y=mu_B, labels="Asset B", pos=4, cex = cex_val)
segments(x0=sig_p[3], y0=0, x1=sig_p[3], y1=mu_p[11], 
         lwd=3, lty="dotted")
text(x=sig_p[3], y=mu_p[3], labels="Inefficient portfolio", 
     pos=4, cex = cex_val)
text(x=sig_p[3], y=mu_p[11], labels="Efficient portfolio", 
     cex = cex_val, pos=4)


```

## portfolios with 1 risky asset and T-Bills. Since the risk free rate fixed over the investment horizon. excess return over thr T-bills or risk premimum. Capital allocation line(CAL). The slope of CAl is called sharpe ration(SR).
```{r}


r_f = 0.03
# T-bills + asset A
x_A = seq(from=0, to=1.4, by=0.1)
mu_p_A = r_f + x_A*(mu_A - r_f)
sig_p_A = x_A*sig_A
sharpe_A = (mu_A - r_f)/sig_A
sharpe_A
# T-bills + asset B
x_B = seq(from=0, to=1.4, by=0.1)
mu_p_B = r_f + x_B*(mu_B - r_f)
sig_p_B = x_B*sig_B
sharpe_B = (mu_B - r_f)/sig_B
sharpe_B


# plot portfolios of T-Bills and assets A and B
plot(sig_p_A, mu_p_A, type="b", col="green", ylim=c(0, max(mu_p_A)),
     xlim=c(0, max(sig_p_A, sig_p_B)), pch=16, cex = cex_val,
     xlab=expression(sigma[p]), ylab=expression(mu[p]), cex_lab = cex_val)
points(sig_p_B, mu_p_B, type="b", col="red", pch=16, cex = cex_val)
text(x=sig_A, y=mu_A, labels="Asset A", pos=4, cex = cex_val)
text(x=sig_B, y=mu_B, labels="Asset B", pos=1, cex = cex_val)
text(x=0, y=r_f, labels=expression(r[f]), pos=2, cex = cex_val)

#
# portfolios of 2 risky assets and T-Bills
#
rho_AB = -0.164
sig_AB = rho_AB*sig_A*sig_B
top = (mu_A - r_f)*sig2_B - (mu_B - r_f)*sig_AB
bot = (mu_A - r_f)*sig2_B + (mu_B - r_f)*sig2_A - (mu_A - r_f + mu_B - r_f)*sig_AB
x_A_tan = top/bot
x_B_tan = 1 - x_A_tan
x_A_tan
x_B_tan
mu_p_tan = x_A_tan*mu_A + x_B_tan*mu_B
sig2_p_tan = x_A_tan^2 * sig2_A + x_B_tan^2 * sig2_B + 2*x_A_tan*x_B_tan*sig_AB
sig_p_tan = sqrt(sig2_p_tan)
mu_p_tan
sig_p_tan

# T-bills plus tangency
x_tan = seq(from=0, to=2.4, by=0.1)
mu_p_tan_tbill = r_f + x_tan*(mu_p_tan - r_f)
sig_p_tan_tbill = x_tan*sig_p_tan

# plot portfolios w/o tangency portfolio
plot(sig_p, mu_p, type="b", pch=16, cex = cex_val,
     ylim=c(0, max(mu_p)), xlim=c(0, max(sig_p)),
     xlab=expression(sigma[p]), ylab=expression(mu[p]), cex_lab = cex_val)
text(x=sig_A, y=mu_A, labels="Asset A", pos=4, cex = cex_val)
text(x=sig_B, y=mu_B, labels="Asset B", pos=4, cex = cex_val)
text(x=sig_p_min, y=mu_p_min, labels="Global min", pos=4, cex = cex_val)
text(x=0, y=r_f, labels=expression(r[f]), pos=2, cex = cex_val)
text(x=sig_p_tan, y=mu_p_tan, labels="Tangency", pos=2, cex = cex_val)
points(sig_p_A, mu_p_A, type="b", col="red", pch=16, cex = cex_val)
points(sig_p_B, mu_p_B, type="b", col="blue", pch=16, cex = cex_val)

# plot portfolios with tangency portfolio
plot(sig_p, mu_p, type="b", pch=16, cex = cex_val,
     ylim=c(0, max(mu_p)), xlim=c(0, max(sig_p)),
     xlab=expression(sigma[p]), ylab=expression(mu[p]), cex_lab = cex_val)
text(x=sig_A, y=mu_A, labels="Asset A", pos=4, cex = cex_val)
text(x=sig_B, y=mu_B, labels="Asset B", pos=4, cex = cex_val)
text(x=sig_p_min, y=mu_p_min, labels="Global min", pos=4, cex = cex_val)
text(x=0, y=r_f, labels=expression(r[f]), pos=2, cex = cex_val)
text(x=sig_p_tan, y=mu_p_tan, labels="Tangency", pos=2, cex = cex_val)
points(sig_p_A, mu_p_A, type="b", col="red", pch=16, cex = cex_val)
points(sig_p_B, mu_p_B, type="b", col="blue", pch=16, cex = cex_val)
points(sig_p_tan_tbill, mu_p_tan_tbill, type="b", col="green", pch=16, cex = cex_val)

#
# interpreting efficient portfolios
#

# safe portfolio: 10% in tangency and 90% in T-bills
mu_safe = r_f + 0.10*(mu_p_tan - r_f)
sig_safe = 0.10*sig_p_tan
# risky portfolio: 110% in tangency and -10% in T-bills
mu_risky = r_f + 2*(mu_p_tan - r_f)
sig_risky = 2*sig_p_tan
plot(sig_p, mu_p, type="b", pch=16, cex = cex_val,
     ylim=c(0, max(mu_p)), xlim=c(0, max(sig_p)), cex_lab = cex_val,
     xlab=expression(sigma[p]), ylab=expression(mu[p]))
text(x=sig_A, y=mu_A, labels="Asset A", pos=4, cex = cex_val)
text(x=sig_B, y=mu_B, labels="Asset B", pos=4, cex = cex_val)
text(x=sig_p_min, y=mu_p_min, labels="Global min", pos=4, cex = cex_val)
text(x=0, y=r_f, labels=expression(r[f]), pos=2, cex = cex_val)
text(x=sig_p_tan, y=mu_p_tan, labels="Tangency", pos=2, cex = cex_val)
points(x=sig_p_tan, y=mu_p_tan, pch=16, cex=2, col="green")
points(sig_p_tan_tbill, mu_p_tan_tbill, type="b", col="green", 
       pch=16, cex = cex_val)
points(x=sig_safe, y=mu_safe, pch=16, cex=2)
text(x=sig_safe, y=mu_safe, labels="Safe", pos=3, cex = cex_val)
points(x=sig_risky, y=mu_risky, pch=16, cex=2)
text(x=sig_risky, y=mu_risky, labels="Risky", pos=3, cex = cex_val)

# find combination of T-bills and tangency that has same SD as asset B
x_t1 = sig_B/sig_p_tan
x_t1
mu_p1_e = r_f + x_t1*(mu_p_tan - r_f)
mu_p1_e
sig_p1_e = x_t1*sig_p_tan
sig_p1_e

# find combination of T-bills and tangency that has same ER as asset B
x_t2 = (mu_B - r_f)/(mu_p_tan - r_f)
x_t2
mu_p2_e = r_f + x_t2*(mu_p_tan - r_f)
mu_p2_e
sig_p2_e = x_t2*sig_p_tan
sig_p2_e

plot(sig_p, mu_p, type="b", pch=16, cex = cex_val,
     ylim=c(0, max(mu_p)), xlim=c(0, max(sig_p)),
     xlab=expression(sigma[p]), ylab=expression(mu[p]))
text(x=sig_A, y=mu_A, labels="Asset A", pos=4, cex = cex_val)
text(x=sig_B, y=mu_B, labels="Asset B", pos=4, cex = cex_val)
text(x=sig_p_min, y=mu_p_min, labels="Global min", pos=4, cex = cex_val)
text(x=0, y=r_f, labels=expression(r[f]), pos=2, cex = cex_val)
text(x=sig_p_tan, y=mu_p_tan, labels="Tangency", pos=4, cex = cex_val)
points(x=sig_p_tan, y=mu_p_tan, pch=16, cex=1_5, col="green", cex = cex_val)
points(sig_p_tan_tbill, mu_p_tan_tbill, type="b", col="green", pch=16, cex = cex_val)
segments(x0=sig_B, y0=0, x1=sig_B, y1=mu_p1_e, lwd=3, lty="dotted")
segments(x0=sig_B, y0=mu_p1_e, x1=0, y1=mu_p1_e, lwd=3, lty="dotted")
segments(x0=sig_p2_e, y0=0, x1=sig_p2_e, y1=mu_p2_e, lwd=3, lty="dotted")
segments(x0=0, y0=mu_p2_e, x1=sig_B, y1=mu_p2_e, lwd=3, lty="dotted")
text(x=sig_B, y=mu_p1_e, labels="e1", pos=3, cex = cex_val)
text(x=sig_p2_e, y=mu_B, labels="e2", pos=3, cex = cex_val)


```

# risk budgeting

```{r}


riskBudgetSD <- function(x_A, x_B, mu_A, mu_B, sig2_A, sig2_B, sig_AB) {
  sig_p = sqrt(x_A^2 * sig2_A + x_B^2 * sig2_B + 2*x_A*x_B*sig_AB)
  mcr_A = (x_A*sig2_A + x_B*sig_AB)/sig_p
  mcr_B = (x_B*sig2_B + x_A*sig_AB)/sig_p
  cr_A = x_A*mcr_A
  cr_B = x_B*mcr_B
  pcr_A = cr_A/sig_p
  pcr_B = cr_B/sig_p
  ans = list(sig_p=sig_p,
             x=c(x_A, x_B),
             mcr = c(mcr_A, mcr_B),
             cr = c(cr_A, cr_B),
             pcr = c(pcr_A, pcr_B))
  return(ans)
}             
x_A = 0.5
x_B = 0.5
rb_sd = riskBudgetSD(x_A, x_B, mu_A, mu_B, sig2_A, sig2_B, sig_AB)             
rb_sd
x_A = 1.5
x_B = -0.5
rb_sd = riskBudgetSD(x_A, x_B, mu_A, mu_B, sig2_A, sig2_B, sig_AB)             
rb_sd

riskBudgetVaR <- function(x_A, x_B, mu_A, mu_B, sig2_A, sig2_B, sig_AB, 
                          w0=100000, alpha=0.05) {
  mu_p = x_A*mu_A + x_B*mu_B
  sig_p = sqrt(x_A^2 * sig2_A + x_B^2 * sig2_B + 2*x_A*x_B*sig_AB)
  VaR_p = (mu_p + sig_p*qnorm(0_05))*w0
  mcr_sig_A = (x_A*sig2_A + x_B*sig_AB)/sig_p
  mcr_sig_B = (x_B*sig2_B + x_A*sig_AB)/sig_p
  mcr_A = (mu_A + mcr_sig_A*qnorm(alpha))*w0
  mcr_B = (mu_B + mcr_sig_B*qnorm(alpha))*w0
  cr_A = x_A*mcr_A
  cr_B = x_B*mcr_B
  pcr_A = cr_A/VaR_p
  pcr_B = cr_B/VaR_p
  ans = list(VaR_p=VaR_p,
             x=c(x_A, x_B),
             mcr = c(mcr_A, mcr_B),
             cr = c(cr_A, cr_B),
             pcr = c(pcr_A, pcr_B))
  return(ans)
}   
x_A = 0.5
x_B = 0.5
rb_VaR = riskBudgetVaR(x_A, x_B, mu_A, mu_B, sig2_A, sig2_B, sig_AB)             
rb_VaR
x_A = 1.5
x_B = -0.5
rb_VaR = riskBudgetVaR(x_A, x_B, mu_A, mu_B, sig2_A, sig2_B, sig_AB)             
rb_VaR
```
